---
title: "Introduction to op3r"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{op3r}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
httptest2::start_vignette("vignette-mocks/op3r")
```

```{r setup}
library(op3r)
```

The `{op3r}` package was created to bring the novel [Open Podcast Prefix Project](https://op3.dev/) (OP3) API to R users through an streamlined and straight-forward interface. A growing number of podcasts adopting the Podcastion 2.0 namespace are opting in to the OP3 service as a robust and transparent source of statistics associated with their podcast.

## Initial setup

The API endpoints supported by `{op3r}` require an API token. To obtain your own set, create a free developer account at <https://op3.dev/api/docs> and save the token as an environment variables called `OP3_API_TOKEN`  within a project-level or default user-directory `.Renviron` file. Below is an example (change to match your token value):

```
OP3_API_TOKEN="abcd123"
```

Once the package is installed, you can verify if authentication is working as expected:

```{r}
library(op3r)
op3_token_isset()
```

## Case Study

The API endpoints offered by the OP3 service offer the following general capabilities:

* Assembling metrics pooled across all podcasts utilizing the OP3 service.
* Metrics associated with a specific podcast using the OP3 service.

The examples below demonstrate each of these perspectives. The podcast used for the single-podcast examples is the [R Weekly Highlights Podcast](https://serve.podhome.fm/r-weekly-highlights) hosted by Eric Nantz and Michael Thomas. First we load additional R packages used for analysis and visualization:

```{r}
library(dplyr)
```


### Podcast metadata

Each podcast using the OP3 project is assigned a universally unique identifier (UUID). Many of the functions tailored to specific podcasts rely on this identifier as a parameter. To find this identifier for a given podcast, you can use either of the following methods:

* Use the `op3_show()` function and supply either the podcast's GUID or RSS feed URL for the required `show_id` parameter: 

```{r}
rweekly_show_rss <- "https://serve.podhome.fm/rss/bb28afcc-137e-5c66-b231-4ffad7979b44"

op3_show(show_id = rweekly_show_rss)
```

* Use the `op3_get_show_uuid()` function with either the podcast GUID or RSS feed URL for `show_id`:

```{r}
op3_get_show_uuid("https://serve.podhome.fm/rss/bb28afcc-137e-5c66-b231-4ffad7979b44")

op3_get_show_uuid("bb28afcc-137e-5c66-b231-4ffad7979b44")
```

### Download Metrics

Within the podcast industry, many hosts and hosting companies gravitate towards download metrics. The `op3_downloads_show()` function obtains download statistics for the previous 30 days of the time the function is executed (for additional details on how OP3 calculates downloads refer to the official [documentation](https://op3.dev/download-calculation)):

```{r}
# using the OP3 UUID for R Weekly Highlights
show_id <- "c008c9c7cfe847dda55cfdde54a22154"
op3_downloads_show(show_id = show_id)
```

If you are interested in the downloads per week captured in the metrics, you can run the same function with `nest_downloads = FALSE`:

```{r}
op3_downloads_show(show_id = show_id, nest_downloads = FALSE) |>
  select(weekNumber, weeklyDownloads)
```

Additional download metrics are available based on applications used to play the podcast. The `op3_too_showsapp()` function obtains metrics over the last three calendar months from the time of execution:

```{r}
op3_top_show_apps(show_id = show_id)
```

```{r, include=FALSE}
httptest2::end_vignette()
```